---
title: "ArbOvl: computing pairwise and multiple genomic interval sets under various conditions and standards"
author: "Angry Programming Nerd"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ArbOvl}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup,results = 'hide',message = FALSE}
library(GenomicRanges)
library(gplots)
library(gtools)
library(UpSetR)
library(ggvenn)
library(ArbOvl)



```
#Description

ArbOvl is a simple wrapper which supports computing pairwise and multiple genomic interval sets intersection under various conditions and standards, the main output is a list containing two objects: the first one is a binary data frame representing intersection sizes in each combination, the second one is a list containing intersecting intervals of each combination. Users can use the binary data frame as input to visualize the result, both venn plot(up to 5 sets) and upset plot are supported. Users can also output the intersecting intervals into a bed-format data frame.

#Introduction

Finding intersection of genomic interval might be the most common scene when manipulating bed files. It can be a little bit more complicated when defining what is overlapping or intersecting. ArbOvl offers various options including minimum overlap bases for defining overlapping, minimum gap length to merge non-disjointing overlapping intervals,etc, which can compute intersection in a much more flexible way. Users can also change the definition of intersection  sizes(e.g. counting number of intersecting bases instead)

```{r}
## loading packages
library(GenomicRanges)
library(IRanges)
library(gplots)
library(gtools)
library(UpSetR)
library(ggvenn)
library(ArbOvl)
```

#Loading bed file into GRanges object
one can load bed files into GRanges objects either by file path or existed data frame

```{r}
file1 <- system.file('extdata','A_peak.bed',package = 'ArbOvl')
gr1 <- bed2gr(file1)
print(gr1)
```

#Loading multiple bed files and arrange them into a list
essential step if you want to perform multiple genomic intervals sets comparison

```{r}
file2 <- system.file('extdata','B_peak.bed',package = 'ArbOvl')
gr2 <- bed2gr(file2)
file3 <- system.file('extdata','C_peak.bed',package = 'ArbOvl')
gr3 <- bed2gr(file3)

grlist <- list('A' = gr1,'B' = gr2, 'C' = gr3)
```

#Compute overlap in default manner
`arbtr_ol` return a list containing two objects: the first one is a binary data frame which can be used as input for visualization; the other one is a list containing intersecting intervals of each combination

```{r}
res <- arbtr_ol(grlist)
print(res[[1]])
print(res[[2]])

```

## Compute overlap under various condition

if one wants to keep intervals with at least n bases overlapping only, applying  a threshold to `minoverlap` in `arbtr_ol` would be the way

```{r}
res <- arbtr_ol(grlist,minoverlap = 3L)
print(res[[1]])
print(res[[2]])
```

if one wants to merge those intersecting intervals which are sufficiently close, apply another threshold to `min.gapwidth` in `arbtr_ol`

```{r}
res <- arbtr_ol(grlist,min.gapwidth = 3L)
print(res[[1]])
print(res[[2]])
```

# Visualize the results

```{r}
res <- arbtr_ol(grlist)

# venn plot
olm2venn(res[['oldf']])

#changing method for visualizing venn plot
olm2venn(res[['oldf']],method = 'ggvenn')

#upset plot
olm2upset(res[['oldf']])
```

# Ouput the intersecting intervals into bed-formatted data frame

```{r}
bed1 <- gr2df(res[['overlap.interval']][2])


#concatenate multiple intervals into one data frame
ttlbed <- gr2df(res[['overlap.interval']])
```

